---
title: "Meta-analytic effect size and publication bias"
author: "Alexandre de Pontes Nobre and Gabriela Melo"
date: "27.03.2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
library(meta)
```


```{r Compute meta-analytic effect size, include=FALSE, cache=TRUE}

# Save defaults
defaults <- par()
options_defaults <- options()

# Read spreadsheet
es_table <- read_delim("C:/Users/Biosig/Google Drive/Doutorado/Tese/Meta-analysis_IB/Dados/Analysis/es_table.txt", 
                          delim="\t", 
                       locale = locale(decimal_mark = "."))


## Estimate correlation between measures (conditions) to compute Cohen's d for variance from two studies:
# Schnuerch et al. (2016)
# Razpurker-apfeld and Pratt (2008)
cor_pairs <- 0.95


# Create variables to compute Hedges' g (formulas from Borestein's Introduction to Meta-Analysis, 2009)

# Compute variance of d (formula 4.28)
es_table$variance_d <- (1/es_table$N_per_group + 
                          (es_table$d)^2/2*es_table$N_per_group) * 2*(1-cor_pairs)

# Compute correction factor J (formula 4.22)
es_table$J <- 1 - (3/(4*(es_table$N_per_group-1)-1))

# Compute Hedges' g (formula 4.23)
es_table$hedgesg <- es_table$J * es_table$d 

# Compute variance of g (formula 4.24)
es_table$variance_g <- (es_table$J)^2 * es_table$variance_d

# Compute standard error of g (formula 4.25)
es_table$se_g <- sqrt(es_table$variance_g)


## Compute meta analytic effect size
meta_es <- metagen(TE = es_table$hedgesg, # treatment effect (Hedge's g)
                   es_table$se_g, #standard error of treatment,
                   studlab = es_table$Study,
                   comb.fixed = TRUE,
                   comb.random = TRUE)

summary(meta_es)
knitr::kable(meta_es)
```

Here are the forest plot and the funnel plot to diagnose publication bias.

```{r Forest plots and funnel plots, echo=FALSE}
# Plots
forest(meta_es)

funnel(meta_es,
       xlab = "Hedges' g")#,
       #studlab = es_table$Study)
```

These plots are highly asymetrical. Here is a test for plot asymetry.

```{r Test for funnel plot asymetry, echo=FALSE}
metabias(meta_es,  method = "rank")
```


The test shows that there is indeed a bias. So, I used the trim-and-fill method, not so much to see how the plot would look like with the inputed dots (it is pretty clear there would be a large dark cloud on the left side) but as to check if the effect would remain after that.

```{r Trim-and-fill method, include=FALSE}
trimmed_meta <- trimfill(meta_es,
                         left = TRUE,
                         ma.fixed = FALSE)
```

```{r Plot trimmed-and-filled funnel plot, echo=FALSE}
funnel(trimmed_meta)

forest(trimmed_meta)
```

Check the analysis results here:

```{r, echo = FALSE}
trimmed_meta
```

So, the effect disappears. In light of this, I suppose we should discuss: 1) checking if this analysis is done correctly; and 2) what the causes for the bias are.
